---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
unzip("activity.zip")
act <- read.csv("activity.csv")
act$date <- strptime(act$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r}
act2 <- na.omit(act)
hist(tapply(act2$steps, as.numeric(act2$date), sum), main = "Total Steps per Day", 
     col = "purple", xlab = "Number of Steps")
meanSteps <- tapply(act2$steps, as.numeric(act2$date), mean)
medianSteps <- tapply(act2$steps[act2>0], as.numeric(act2$date[act2>0]), 
        median, na.rm = TRUE)
meanMedian <- cbind(rownames(table(as.character(act2$date))), meanSteps, medianSteps)
meanMedian
```

## What is the average daily activity pattern?

```{r}
meanInterval <- cbind(rownames(table(act2$interval)), tapply(act2$steps, 
        act2$interval, mean))
plot(meanInterval[,1], meanInterval[,2], type = "l", col = "blue", xlab = 
        "Interval", ylab = "Average Steps per Day", 
        main = "Average Steps by Time of Day")
maxSteps <- c(subset(meanInterval, meanInterval[,2] ==
        max(as.numeric(meanInterval[,2])), select = 
        meanInterval[,1]:meanInterval[,2]), max(as.numeric(meanInterval[,2])))
maxSteps
```

## Imputing missing values

```{r}
sum(is.na(act))
act3 <- act
for (i in which(is.na(act3))){
        act3[i,1] <- meanSteps[((i-1)%%288)+1]
}
hist(tapply(act3$steps, as.numeric(act3$date), sum), col = "red", 
     main = "Sum of Steps per Day", xlab = "Steps Taken")
meanSteps2 <- tapply(act3$steps, as.numeric(act3$date), mean)
medianSteps2 <- tapply(act3$steps[act2>0], as.numeric(act3$date[act2>0]), 
        median, na.rm = TRUE) # THIS IS BROKEN
meanMedian2 <- cbind(rownames(table(as.character(act3$date))), meanSteps2, 
        medianSteps2)
meanMedian2
mean(act3$steps, na.rm = TRUE)
median(act3$steps, na.rm = TRUE) # THIS IS BROKEN
mean(act3$steps, na.rm = TRUE) - mean(act$steps, na.rm = TRUE)
median(act3$steps[act2>0], na.rm = TRUE) - median(act$steps[act2>0], na.rm = TRUE)
```

The mean and median from the original dataset and that with averaged NA values
is identical.  Adding mean NA values did, however, increase the daily sum of
steps taken.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
require(ggplot2)
weekend <- c("Saturday", "Sunday")
act[, "weekday"] <- ifelse(weekdays(act$date) == weekend, "weekend", "weekday")
avg <- aggregate(steps ~ interval + weekday, data = act, mean)
ggplot(avg, aes(interval, steps)) +
        geom_line() + facet_grid(weekday ~.)
```
